\documentclass[11pt]{article}

% ---------------------------------------------------------------------------
% Earned Autonomy --- an engineering-leader whitepaper on the 2026-07-28 MCP
% spec. Clean article class (no venue branding). Root owns preamble + metadata;
% one \input per section. Build: cd paper && latexmk -pdf main.tex
% ---------------------------------------------------------------------------

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage[numbers,sort&compress]{natbib}
\usepackage[hyphens]{url}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=blue]{hyperref}

% tighten section spacing to keep the page budget
\titlespacing*{\section}{0pt}{1.4ex plus 0.3ex minus 0.2ex}{0.9ex}

% listings style for the illustrative policy exhibit
\lstdefinestyle{policy}{%
  basicstyle=\ttfamily\small,
  frame=single,
  framerule=0.4pt,
  breaklines=true,
  showstringspaces=false,
  columns=fullflexible,
  keepspaces=true,
  xleftmargin=0.8em,
  xrightmargin=0.8em,
  aboveskip=0.6em,
  belowskip=0.4em,
}

\renewcommand{\abstractname}{Executive summary}

\title{\vspace{-1.2em}\textbf{Earned Autonomy}\\[0.35em]
  \large Conditioning Agent Trust on Delivered Outcomes, Not Task Accuracy}
\author{Andrew Zigler}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
\noindent Enterprises have adopted autonomous software agents faster than they
have learned to tell whether the agents are delivering. The dominant answer to
that trust problem, published under names like earned, graduated, and
progressive autonomy, conditions an agent's authority on how accurately it
performs its tasks. This paper argues that task accuracy is the wrong adaptive
signal. Autonomy conditioned on task accuracy re-creates at the trust layer the
same clean-record trap the audit layer already has: a fully authorized,
correctly typed, cleanly traced call can open a defective change while every
record reads clean, and an accuracy-graded ladder promotes a workflow at the
moment its output draws the least scrutiny. Delivered outcome, whether an
agent's change held in production, is the adaptive conditioning variable these
autonomy ladders leave out. It complements the preventive gates that screen
every change regardless of a workflow's record; it does not replace them. The
Model Context Protocol revision finalizing on 2026-07-28 is the first open
protocol standard to make per-action agent-tool authorization a first-class,
session-independent property, so a workflow's earned tier can be enforced per
action while its outcomes are evaluated on a cadence. The paper walks what the
release makes legible and enforceable, defines a \emph{delivery evidence layer}
that joins governed actions to their delivered outcomes, and describes earned
autonomy: a rolling trust cadence, modeled on the standard's own governance by
measurement, that grants and revokes agent authority on evidence.
\end{abstract}

\input{sections/01-question}
\input{sections/02-legible}
\input{sections/03-record}
\input{sections/04-layer}
\input{sections/05-earned-autonomy}
\input{sections/06-recommendations}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
