\section{What the spec makes legible and enforceable}

After the 2026-07-28 revision, every agent action is at once a well-formed record
and a policy decision point. That, not the size of the release, is what makes it
matter for a governance argument. Five properties produce that result. None is
invented here; each is a mechanism the release already ships, and the point is
what they add up to.

\paragraph{Identity-bound.} Every request now carries a client label in its
\texttt{\_meta} block rather than establishing it once per session: the
\texttt{io.modelcontextprotocol/clientInfo} field travels with the call itself
(the Specification Enhancement Proposal SEP-2575) \citep{mcpchangelog}. That
label is self-declared by the client, not authenticated. The authenticated
principal comes from the authorization side: the release hardens OAuth so a
client must validate a present issuer (\texttt{iss}) parameter against the
recorded authorization server before it redeems a code, closing a class of
mix-up attack (SEP-2468) \citep{mcpchangelog}. An action arrives attributable to
an OAuth-authenticated principal, with \texttt{clientInfo} adding a
client-asserted implementation label that is useful for attribution only when it
is bound to an issued credential.

\paragraph{Typed.} Tool inputs and outputs are lifted to full JSON Schema
2020-12 (SEP-2106), so the shape of every call and its result is declared and
checkable rather than conventional \citep{mcpchangelog}. A record whose payloads
validate against a schema is a record a downstream system can parse without
guessing at its structure.

\paragraph{Traceable.} The release documents W3C Trace Context conventions
(\texttt{traceparent}, \texttt{tracestate}, and \texttt{baggage} in
\texttt{\_meta}, per SEP-414), so a single action correlates across SDKs,
gateways, and any OpenTelemetry backend \citep{mcpchangelog}. The deprecation of
the old Logging primitive points the same way: diagnostic logging moves to
\texttt{stderr} and OpenTelemetry (SEP-2577) \citep{deprecated}. Observability
belongs to the tracing plane, not to a bespoke protocol message.

\paragraph{Enforceable.} Because the release removes protocol sessions
(SEP-2567), any request can land on any server replica, and authorization is
therefore evaluated on every call rather than inherited from a session
established once \citep{sep2567}. The release also requires routing headers,
\texttt{Mcp-Method} and \texttt{Mcp-Name}, on Streamable-HTTP POSTs (SEP-2243)
\citep{mcpchangelog}, which lets a gateway reject a request whose headers and
body disagree. A gateway can now make an allow, deny, or downgrade decision on
the operation from the headers alone. The per-call boundary is a real
enforcement point.

\paragraph{Fleet-scalable.} List and read results now carry required cache
metadata: a \texttt{ttlMs} freshness hint and a \texttt{cacheScope} of
\texttt{public} or \texttt{private} (SEP-2549) \citep{mcpchangelog}. With
sessions gone, these results are stable enough to cache, which makes the served
context and tool listings cheap to reuse at fleet scale. Authorization stays
per-call; caching lowers the data-fetch cost, not the cost of the decision made
on every request.

Set these five properties side by side and the shape is clear. An organization
running agents on the 2026-07-28 spec gets, from the protocol alone, a stream of
actions that are attributable, typed, traceable, individually enforceable, and
cheap to observe at fleet scale. This is a genuine advance in what a protocol
makes governable, and the temptation is to stop here and call the governance
problem solved. It is not solved. A well-formed record has a limit, and that
limit is the subject of the next section.
