\section{What the record cannot tell you}

A record can be complete and still be silent about the thing that matters.
Consider an agent action that is fully authorized, correctly typed, cleanly
traced, and schema-valid, a record with nothing wrong in it, that opens a pull
request introducing a SQL injection. Nothing in the protocol layer is equipped to
notice, because nothing in the protocol layer looks at what the change does once
it is merged. Veracode's 2025 analysis found that 45\% of AI-generated code
samples introduced an OWASP Top-10 vulnerability, across more than a hundred
models, eighty tasks, and four languages, with the rate roughly flat regardless
of model size or recency \citep{veracode2025}. The defective change and the
correct change produce identical protocol records.

This is the clean-record trap, and the audit layer has always had it. A log
tells you an action happened and was permitted; it never tells you the action
was any good. Conditioning autonomy on task accuracy re-creates that trap one
level up, because task accuracy measures the agent's behavior at the call and
sees nothing of what the call delivers. A workflow that scores 99\% on tool-call
correctness is graded on whether its calls are well-formed, not on what they
deliver: the defective call \emph{is} a correct call. Worse, an autonomy ladder
built on task accuracy promotes a workflow precisely when its accuracy scores are
high, which is the moment its output is drawing the least human scrutiny. The
metric rises as the examination falls.

What lies past the call divides into two failure classes, and a division of labor
Section 5 returns to. Latent defects, the security, privacy, and license problems
the SQL injection stands for, ride clean through every production signal and are
caught only by reading the change itself; they belong to the preventive gates.
Delivery failures, the reverts, rework, and changes that sit unmerged, are caught
only by watching what shipped, and that is this paper's subject. An industry
benchmark of 2.7 million pull requests across 253 organizations supplies the
field evidence: the output that matters is delivered work, not generated code,
and agent-opened pull requests merge at markedly lower rates than human-opened
ones, because a change no human owns tends to sit unmerged \citep{aibench2026}.
The variable that counts lives one step past the tool boundary, at the delivered
change, and no protocol call runs there.
