\section{Earned autonomy}

The 2026-07-28 release ships into a conformance-gated, SDK-tiered ecosystem: a
conformance suite live since 2026-01-23, published SDK tiers since 2026-02-23,
and, as of SEP-2484 (Final 2026-06-04), a standing rule that future
Standards-Track SEPs cannot reach Final without a matching conformance scenario
\citep{sep2484, sdktiers}. The SDK tiers are the part worth studying. A Tier 1
SDK must pass 100\% of the conformance suite; if any test fails continuously for
four weeks, it is relegated to Tier 2; and features move through a lifecycle with
a minimum twelve-month deprecation window measured from the revision release
\citep{sdktiers, mcplifecycle}. Standing is earned by demonstrated conformance
and lost publicly when the demonstration lapses.

That is a working model of earned standing, and humans run it: no automated
controller relegates an SDK. Maintainers publish tiers against measured
conformance, on defined clocks, with a defined relegation rule, and the
governance works. This is the precise shape to port. Earned autonomy is a rolling
trust review in which workflow configurations hold published autonomy tiers,
delivery-evidence-layer signals are the input, promotion and relegation follow
defined triggers, and the resulting grant is enforced per call at the boundary
the spec provides. Steer is that cadence: a review rhythm with a clock and a
rule, of the same kind the standard already runs on itself. The closed-loop
controller that reads a metric and rewrites policy is the forward agenda, not the
claim here.

Every component of this cadence is mechanically expressible today: identity and
per-call enforcement from the spec, attribution and outcome evaluation from the
delivery evidence layer, published tiers and clocks from process. The only
genuinely new input is the swap at the center: replace static scope grants,
issued once and rarely revisited, with grants conditioned on delivered-outcome
evidence.

Stated plainly: autonomy conditioned on task accuracy re-creates at the trust
layer the same clean-record trap the audit layer already has. Delivered outcome
is the adaptive conditioning variable every published ladder leaves out, and the
2026-07-28 spec is the first open protocol standard (proprietary gateways
enforced per-call before it) to make per-action agent-tool authorization a
first-class, session-independent property. The grant a workflow has earned is
enforced per action; its delivered outcome is evaluated on a cadence.

Earned autonomy is one of two layers, and it is the adaptive one. Beneath it
sits a preventive layer that screens every change before merge regardless of
which workflow authored it or what record that workflow holds: pre-merge
security scanning, blast-radius and impact gating, and agentic code review. A
cadence tuned to delivered outcomes is blind to the latent-defect classes
(security, privacy, license) that ride clean through every production signal
until they are exploited. The SQL injection of the earlier section holds in
production, passes each outcome check, and would earn a workflow more scope on
outcome evidence alone. Only a gate that reads the change itself catches it.
Earned autonomy governs how much a workflow is trusted to act unreviewed over
time; it presupposes, and does not replace, the review of the individual change.

\paragraph{The failure modes are dials, not objections.} Because Steer is a
cadence, its known failure modes are design parameters that set the cadence's
dials. \emph{Attribution lag} (outcomes arrive days or weeks after the change)
sets the length of the trailing window the review scores on; instantaneous
scoring is the mistake the window length corrects. \emph{Unstable identity}
(agent instances are ephemeral and interchangeable) sets the unit of trust: the
durable thing to govern is a workflow configuration, not an instance, because the
configuration is what produced the work and what will produce it again.
\emph{Goodhart's law} sets the gating signal: gate on held-in-production outcomes
and never on activity counts, because the moment the metric is a throughput count
it is gamed. \emph{Small-n} (too few observations to trust) sets the
cold-start default: absent evidence, a workflow starts in a sandbox tier, since
autonomy is earned from evidence and the safe default without it is low.

\paragraph{Two clarifications keep the framework honest} about what the protocol
does and does not provide. First, the per-call boundary is an enforcement point,
not a scope engine. The spec gives the point at which authorization is checked on
every call; it does not decide what a given principal may do. The scope
semantics, the policy itself, remain the implementer's design. Authentication is
not authorization scope. Second, per-call identity attributes to the
authenticated principal and the named client implementation, carried in
\texttt{clientInfo}, and not to an individual agent or subagent. Agent-level
attribution is a convention an organization builds on top, and it is exactly why
the unit of trust is a workflow configuration: that is the granularity the
evidence layer can attribute to, not a per-instance identity the protocol never
reported.

This shape has precedent outside software delivery. The CNCF moves a project from
Sandbox to Incubating to Graduated on demonstrated adoption, security review, and
governance maturity, ratified by humans rather than by a metric gate
\citep{cncflifecycle}. That is the posture earned autonomy takes toward a
workflow: evidence-based maturation with human ratification.

There is also direct empirical support for building the feedback capability the
framework depends on. DORA's 2025 research found that AI acts as an amplifier: it
has a positive relationship with delivery throughput and a negative one with
stability, and which effect dominates is conditional on a team's existing
capabilities \citep{dora2025}. Among those capabilities is the delivery-feedback
capability this paper argues an organization must build. The amplifier result
corroborates that delivery stability is the thing at risk. It motivates caring
about outcomes; it does not, by itself, pick the conditioning variable.

The cadence is configuration, not research. The trust decision, reading the
delivered-outcome evidence and setting the tier, is the organization's to make;
the gateway's only job is to enforce the resulting scope on every call. The
gateway never consults evidence at call time: the cadence compiles evidence into
an identity assignment, and the gateway enforces that identity. A workflow holds
the identity its current tier has earned, and the gateway gates each tool call
against that identity. Because a tier is bound to a credential, relegation is not
complete until the prior tier's credential is revoked, not merely superseded: a
workflow that keeps a live higher-tier key keeps the scope it was supposed to
lose. Issuing, rotating, and revoking those credentials is the organization's
responsibility, not the protocol's.

\begin{figure}[!ht]
\centering
\begin{minipage}{0.92\linewidth}
\begin{lstlisting}[style=policy]
# Per-call authorization at an MCP gateway. The cadence maps each
# workflow to a tier identity at credential issuance; the gateway
# enforces that identity's scope on every tool call. Relegation =
# the cadence reissues the workflow a lower-tier identity (no edit
# to this policy), and its next call is gated anew.
authorization:
  - allow: identity == "tier-trusted"       # full access
  - allow: identity == "tier-sandbox"       # read-only
        && tool in ["read_file", "list_directory", "search_files"]
\end{lstlisting}
\end{minipage}
\caption{\textbf{Actuator validated on agentgateway}, an open-source MCP gateway.
A per-call authorization rule gates tool scope on a workflow's tier-earned
identity; the identical \texttt{write\_file} call was allowed for the trusted
identity and denied, per call, for the sandbox identity, because
\texttt{write\_file} is absent from that tier's allowlist, while the read call was
allowed for both (run notes in \texttt{data/}). This validates the actuator, the
per-call identity-scoped authorization point; the sensor (the delivery evidence
layer) and the policy (the cadence) are described in this paper, not demonstrated
here. The pattern needs only per-call, identity-scoped policy and runs on any
gateway with that capability. The evidence-to-tier decision is the organization's
cadence; the gateway is only the actuator, and the unit of trust is a workflow
configuration, not an instance. It is the shape teams already use to gate a
deploy, pointed at delivered-outcome evidence.}
\end{figure}
